Learning by Teaching a Robot:\\The Case of Handwriting

### 2016-01-13 rebuttal.tex

\section*{Reporting of the studies}

We have clarified our wording to only draw conclusion from factual measured
data or feedback from the parents/teachers/therapists (this is in particular
the case for the case-study 1),

We have added quantitative data in the reporting of the case study 2, that
explicitly show a quantifiable improvement of handwriting. We want however to remain
cautious in our claims: handwriting acquisition is a complex cognitive skill.
While the child demonstrated clear improvements over the course of our
experiment, measuring the lasting effects of the robot-based remediation is
difficult (as many other individual and social factors may have impacted the
child's performance as well).

We have also added the results of an additional experiment (conducted during the
summer 2015), that involved 8 children with handwriting deficits. This
experiment focused on assessing (teacher) role adoption by the children when 
teaching the robot.

Finally, we have added new sections (sections 1.3 and 4) that respectively
explicitly state our research questions, and summarizes and discusses our findings.

### 2016-01-13 main.tex
Proof-reading intro

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{henry}
    \caption{\small Thomas teaching a \sc{Nao} robot how to write numbers, with 
    the help of an occupational therapist.}
    \label{fig:henry}
\end{figure}

@@

\paragraph{Acceptance} A second research question that we study here relates to
the acceptance of such a system in educative and medical contexts. Remediation
to cognitive deficits is a sensitive question for the caretakers, all the more
so when working with children. Introducing a robot-based system in such
situations is not self-evident, especially as the mentor-protégé relationship
further relies on building strong affective and cognitive bonds between the
child and the robot (\ie the robot is not simply experienced as a passive tool,
as would be a computer program for instance). By establishing collaborations
with teachers, parents and therapists, we aim at going to the practitioners'
field level and explore the conditions for such a robot-based system to be
accepted.

\paragraph{Engagement} We have already mentioned that handwriting remediation
requires exercising over long periods of time. In terms of human-robot
interaction, this translates into a third research question: how to sustain
several hours of interaction while performing a task (handwriting) which is
essentially a boring routine? This is usually studied in term of
\emph{engagement}: how to engage, and thereafter sustain the engagement, of a
child into an interaction with a robot? As already hinted, we hypothesize that
the \emph{learning by teaching} paradigm may prove an effective lever to get the
child to commit to help the robot over an extended period of time. This needs to
be evidenced, which incidentally calls for actually running interaction
experiments over long durations (several hours spread over several weeks).

\paragraph{Remediation Efficacy} Finally, the last, key research question is: is
our approach effective in helping children to recover from handwriting deficits?
Answering this question in a broad, general way is unlikely as {\it 1)} the
observed deficits vary a lot between the children (no single metric) and have
complex, compound causes, and, {\it 2)} ``learning how to write'' is not a
monolithic, linear skill that could be easily rated from 0 to 1.  However, by
crossing qualitative assessments made by the caretakers with quantitative
measures of visual similarity between the letters drawn by the children and some
reference templates, we aim at assessing the impact of our system on the
children' handwriting, at least in terms of trend.

@@

We then present and report on the field experiments that we have conducted over
the last two years, including four studies at schools
(Section~\ref{studies_school}), one longer experiment with
eight children in an occupational therapy clinic (Section~\ref{study_normandy}), and two one-month
long case studies (Sections~\ref{vincent} and~\ref{thomas}). While the focus of the school experiments was mostly the
technical validation and data acquisition, the three other experiments involved
children with actual deficits, and gave us initial insights into the relevance and
effectiveness of our approach.

>>>>

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{henry}
    \caption{\small Thomas teaching a {\sc Nao} robot how to write numbers, with 
    the help of an occupational therapist.}
    \label{fig:henry}
\end{figure}

@@

\paragraph{Acceptance} A second research question that we study here relates to
the acceptance of such a system in educative and medical contexts. Remediation
to cognitive deficits, even mild ones, is a sensitive question for the
caretakers, all the more so when working with children. Introducing a
robot-based system in such situations is not self-evident, especially as the
mentor-protégé relationship further relies on building strong affective and
cognitive bonds between the child and a machine (here, the robot is not simply
experienced as a passive tool, as would be an educational computer program for
instance: it instead plays a potentially disruptive role). By establishing
collaborations with teachers, parents and therapists, we aim however at going
down to the practitioners' field level and explore the conditions for such a
robot-based system to be accepted.

\paragraph{Engagement} We have already mentioned that handwriting remediation
requires exercising over long periods of time. In terms of human-robot
interaction, this translates into a third research question: how to sustain
several hours of interaction while performing a task (handwriting) which is
essentially perceived as a boring routine? Sustained interaction is usually
studied in term of \emph{engagement}: how to engage, and thereafter sustain the
engagement, of a child into an interaction with a robot? We hypothesize that the
\emph{learning by teaching} paradigm may be an effective lever to get the child
to commit in helping the robot over an extended period of time. This needs to be
evidenced -- which incidentally calls for setting up and carrying long-term
child-robot experiments.

\paragraph{Remediation Efficacy} Finally, the last, key, research question is: is
our approach effective in helping children to recover from handwriting deficits?
Answering this question in a broad, general way is unlikely as {\it 1)} the
observed deficits vary a lot between the children (no single metric) and have
complex, compound causes, and, {\it 2)} ``learning how to write'' is not a
monolithic, linear skill that could be easily rated from 0 to 1.  However, by
crossing qualitative assessments made by the caretakers with quantitative
measures of visual similarity between the letters drawn by the children and some
reference templates, we aim at assessing the impact of our system on the
children' handwriting, if only in terms of trends.

@@

We then present and report on the field experiments that we have conducted over
the last two years, including four studies at schools
(Section~\ref{studies_school}), one longer experiment with eight children in an
occupational therapy clinic (Section~\ref{study_normandy}), and two one-month
long case studies (Sections~\ref{vincent} and~\ref{thomas}). While the focus of
the school experiments was mostly the technical validation of the system and
data acquisition, the three other experiments involved children with actual
deficits, and gave us initial insights into the relevance and effectiveness of
our approach.
